### 2022-11-17 lmcs/conclusion.tex
clarifying some related work

While our work focuses on \emph{binary} session types, it is worth to discuss related work on Multiparty Session Types (MPST). The line of work on MPST starts with Honda~\etal~\cite{hondayoshida08}, which guarantees deadlock freedom \emph{within a single session}, but not for session interleaving.
Bettini~\etal~\cite{bettinicoppo08} follow a technique similar to Kobayashi's for MPST.
The main difference with our work is that we associate priorities with communication actions, where Bettini~\etal~\cite{bettinicoppo08} associate them with channels.
Carbone and Montesi~\cite{carbonemontesi13} combine MPST with choreographies and obtain a formalism that satisfies deadlock freedom.
Deni\'{e}lou and Yoshida~\cite{DenielouY13} introduce \emph{multiparty compatibility} which generalises duality in binary session types. They synthesise safe and deadlock-free global types from local types leveraging LTSs and communicating automata.
Castellani~\etal~\cite{CastellaniDGH20} guarantee lock freedom, a stronger property than deadlock freedom, for MPST with \emph{internal delegation}, where participants in the same session are allowed to delegate tasks to each other, and internal delegation is captured by the global type.
Scalas and Yoshida~\cite{scalasyoshida19} provide a revision of the foundations for MPST, and offer a less complicated and more general theory, by removing duality/consistency. The type systems is parametric and type checking is decidable, but allows for a novel integration of model checking techniques. More protocols and processes can be typed and are guaranteed to be free of deadlocks.

>>>>

While our work focuses on \emph{binary} session types, it is worth to discuss related work on Multiparty Session Types (MPST), which describe communication among multiple agents in a distributed setting. The line of work on MPST starts with Honda~\etal~\cite{hondayoshida08}, which guarantees deadlock freedom \emph{within a single session}, but not for several session interleaving.
Bettini~\etal~\cite{bettinicoppo08} follow a technique similar to Kobayashi's for MPST.
The main difference with our work is that we associate priorities with communication actions, where Bettini~\etal~\cite{bettinicoppo08} associate them with channels.
Carbone and Montesi~\cite{carbonemontesi13} combine MPST with choreographic programming and obtain a formalism that satisfies deadlock freedom. In the same vein as MPST, choreographic programming specifies communication among all participants in a distributed system. While MPST target mainly protocol descriptions, choreographies have mainly targetted implementations and programming languages as they are suitable for describing concrete system implementations.
Deni\'{e}lou and Yoshida~\cite{DenielouY13} introduce \emph{multiparty compatibility}, which generalises the notion of duality in binary session types. They synthesise safe and deadlock-free \emph{global types}--specifying communication among \emph{all} involved participants, from \emph{local session types}--specifying communication from the viewpoint of \emph{one} participant. To do so, they leverage Labelled Transition Systems (LTSs) and communicating automata.
Castellani~\etal~\cite{CastellaniDGH20} guarantee lock freedom, a stronger property than deadlock freedom, for MPST with \emph{internal delegation}, where participants in the same session are allowed to delegate tasks to each other, and internal delegation is captured by the global type.
Scalas and Yoshida~\cite{scalasyoshida19} provide a revision of the foundations for MPST, and offer a less complicated and more general theory, by removing duality/consistency. The type systems is parametric and type checking is decidable, but allows for a novel integration of model checking techniques. More protocols and processes can be typed and are guaranteed to be free of deadlocks.
